### 2015-05-15 evaluation.tex
fixes

Most of the races that \whoop and \corral discovered can be classified in two cases. The first case is accessing a global counter (or flag) from concurrent entry points, without holding a lock. This might be for performance, and indeed a lot of the races we found might be benign. Even benign races, though, lead to undefined behavior according to the C standard, and it is well known that undefined behaviors can lead to unexpected results when combined with aggressive compiler optimizations. The second case is an entry point modifying a field of a struct (either global or passed as a parameter) without holding a lock. This can lead to a race if another entry point simultaneously accesses the same field of the same struct.

As an example of the second case, we found the following race in the generic\_nvram driver (see Figure~\ref{fig:data_race_example}): in the \texttt{llseek} entry point the driver accesses the file offset \texttt{file->f\_pos} without a lock (\texttt{file} is passed as a parameter to \texttt{llseek}). This causes a race if the driver invokes \texttt{llseek} from another thread passing the same \texttt{file} object as a parameter.  We observed that another char driver, using the same APIs, \emph{does} use a mutex to protect the offset access in its \texttt{llseek} entry point, leading us to suspect that the race we found in generic\_nvram is not benign.  We have filed a bug report in the Linux kernel bug tracker and are awaiting a response.

>>>>

Most of the races that \whoop and \corral discovered fall into two categories. The first is about accessing a global counter (or flag) from concurrent entry points, without holding a lock. This might be for performance, and indeed a lot of the races we found might be benign. Even benign races, though, lead to undefined behavior according to the C standard, and it is well known that undefined behaviors can lead to unexpected results when combined with aggressive compiler optimizations. The second is about an entry point modifying a field of a struct (either global or passed as a parameter) without holding a lock. This can lead to a race if another entry point simultaneously accesses the same field of the same struct.

As an example of the second category, we found the following race in the generic\_nvram driver (see Figure~\ref{fig:data_race_example}): the \texttt{llseek} entry point accesses the file offset \texttt{file->f\_pos} without a lock (\texttt{file} is passed as a parameter to \texttt{llseek}). This causes a race if the driver invokes \texttt{llseek} from another thread passing the same \texttt{file} object as a parameter.  We observed that another char driver, using the same APIs, \emph{does} use a mutex to protect the offset access in its \texttt{llseek} entry point, leading us to suspect that the race we found in generic\_nvram is not benign.  We have filed a bug report in the Linux kernel bug tracker and are awaiting a response.
